---
title: "RQ2"
output: html_notebook
---

# Requirements

### Libraries
```{r, message=F}
library(tidyverse)
library(cluster)
library(mclust)
library(gridExtra)
library(mongolite)
library(factoextra)
library(gridExtra)
library(fclust)
library(stringr)
library(flextable)
```


### Database connection
```{r}
connection_string <- 'mongodb://192.168.1.114:27018'
collection_name <- "commits"
database_name <- "epm-test-smell"
commits_collection <- mongo(collection=collection_name, db=database_name, url=connection_string)
```

# Functions
```{r}
format_users <- function(data){
  data %>% select(internal, external) %>% rename(Internal = internal, External = external) %>% gather(key = Metric, value= Expertise) %>%  group_by(Metric, Expertise) %>% summarise(count = n()) %>% mutate(Proportion = count/sum(count))
}

plot_users <- function(data){
  format_users(data) %>% ggplot(aes(x=Metric, y=Proportion,fill = Expertise, label = scales::percent(Proportion))) + geom_bar(stat = 'identity',position = position_dodge()) +     geom_text(position = position_dodge(width = .9),vjust = -0.5, size = 3) +
  scale_fill_grey(start = 0.5, end = 0.7, na.value = "red") +
  theme_minimal() + theme(axis.title = element_text(), legend.position = "bottom")

}

plot_by_types <- function(data){
 data %>% ggplot(aes(y = smells, x = percentage, fill=Expertise, label = scales::percent(percentage))) +
  geom_bar(stat = "identity", position="dodge") +
  geom_text(aes(label =scales::percent(percentage, accuracy = 0.1), hjust=+0.5), position = position_dodge(width = .8), size = 3) +
  scale_fill_grey(start = 0.5, end = 0.7, na.value = "red") + labs(x ="Proportion", y = "Smell Type") + theme_minimal() + theme(axis.title = element_text(), legend.position = "bottom")
}

save_plot <- function(p,name=deparse(substitute(p)), width = 13, heigth = 11){
ggsave(p, file=paste('images/',name,'.pdf' ,sep=''), device = "pdf", width = width, height = heigth, units = 'cm')
}

execute_test <- function(d1, d2){
  fisher.test(t(rbind(table(d1), table(d2))))
}
```

# RQ2 What is the impact of developper expertise on test smells

## Summary

```{r}
#Get total github users count
total_users <- commits_collection$aggregate('
[
  {
    "$match":{
      "url":{
        "$regex":"github",
        "$options":"i"
      }
    }
  },
  {
      "$set": {
          "users": {
              "$objectToArray": "$users"
          }
      }
  }, 
  {
    "$group":{
      "_id":"$users.v.github_id"
    }
  },
  {
    "$count":"Github users"
  }
]
')

#Get total project count
total_projects <- commits_collection$aggregate('
[
  {
    "$match":{
      "url":{
        "$regex":"github",
        "$options":"i"
      }
    }
  },
  {
    "$group":{
      "_id":"$app"
    }
  },
  {
    "$count":"Projects"
  }
]
')

total_commits <- commits_collection$aggregate('
[
  {
    "$match":{
      "url":{
        "$regex":"github",
        "$options":"i"
      }
    }
  },
  {
    "$count":"Commits"
  }
]
')

totals <- cbind(total_projects, total_commits, total_users) 

```

```{r}
flextable(totals) %>% colformat_int(big.mark = ' ') %>% autofit()

```

## Clustering

### Data preparation
```{r}
#Get all users
users <- commits_collection$aggregate('
  [
    {
        "$set": {
            "users": {
                "$objectToArray": "$users"
            }
        }
    }, {
        "$unwind": "$users"
    }, {
        "$match": {
            "users.v.type": "User"
        }
    }, {
        "$group": {
            "_id": "$users.v.github_id", 
            "nCommits": {
                "$sum": 1
            }, 
            "nCommitsAuthor": {
                "$sum": {
                    "$cond": [
                        {
                            "$eq": [
                                "$users.k", "author"
                            ]
                        }, 1, 0
                    ]
                }
            }, 
            "nCommitsCommitter": {
                "$sum": {
                    "$cond": [
                        {
                            "$eq": [
                                "$users.k", "committer"
                            ]
                        }, 1, 0
                    ]
                }
            }, 
            "user": {
                "$last": "$users.v"
            }
        }
    }, {
        "$project": {
            "user.github_id": 0
        }
    }, {
        "$sort": {
            "nCommits": -1
        }
    }
]
')

users <- do.call(cbind,users)
#Remove application user
users <- users %>% filter(user.login != 'web-flow' & user.login != 'weblate') %>% select(nCommitsCommitter, nCommitsAuthor, user.followers, user.public_gists, user.public_repos, user.login,'_id')
users <- na.omit(users)
```

```{r}
# select important fields for expertise outside the projects
users_external_info <- users %>% select(user.followers, user.public_gists, user.public_repos)

#Convert dates to timestamps
#users_external_info <- users_external_info %>% mutate(user.created_at = as.numeric(as.POSIXct(user.created_at)), user.updated_at = as.numeric(as.POSIXct(user.updated_at)))

#Scale fields
users_external_info <- scale(users_external_info)
head(users_external_info)
```


```{r}
#Select important fields for expertise inside the projects
users_internal_info <- users %>% select(nCommitsAuthor, nCommitsCommitter)
#Scale fields
users_internal_info <- scale(users_internal_info)
head(users_internal_info)
```

### K-means
```{r}
kmeans2_internal <- kmeans(users_internal_info, centers = 2, nstart = 25)
kmeans2_external <- kmeans(users_external_info, centers = 2, nstart = 25)

#Set user cluster
users$internal <- as.factor(kmeans2_internal$cluster)
users$external <- as.factor(kmeans2_external$cluster)

#Use the user kevinsawicki as reference for senior programmer
internal_senior_cluster_label <- kmeans2_internal$cluster[users$user.login == 'kevinsawicki']
external_senior_cluster_label <- kmeans2_external$cluster[users$user.login == 'kevinsawicki']

expertise_labels <- as.factor(c('Junior','Senior'))

#Update users expertise label
users <- users %>% mutate(internal = case_when(
  internal == internal_senior_cluster_label ~ expertise_labels[2],
  internal != internal_senior_cluster_label ~ expertise_labels[1],
))

users <- users %>% mutate(external = case_when(
  external == external_senior_cluster_label ~ expertise_labels[2],
  external != external_senior_cluster_label ~ expertise_labels[1],
))

#Compare the two clusters
print(adjustedRandIndex(users$internal, users$external))
```


```{r}
p_cluster_internal <- fviz_cluster(kmeans2_internal, data=users_internal_info, axes = c(0,1), xlab = 'Scaled number of commits (author)', ylab = 'Scaled number of commits (committer)', main='') + theme_minimal() + scale_fill_grey(start = 0.5, end = 0.7, na.value = "red")
p_cluster_external <- fviz_cluster(kmeans2_external, data=users_external_info, axes = c(0,1)) + scale_fill_grey(start = 0.5, end = 0.7, na.value = "red") + theme_minimal()

save_plot(p_cluster_internal)
print(p_cluster_internal)

```

```{r}

p_users <- plot_users(users)
save_plot(p_users)
p_users

```

## RQ2.1 How does the distribution of developper updating test files compare to the contributors distribution

```{r}
#Get all unique users with at least one commit impacting a test file.
user_changing_test_files <- commits_collection$aggregate('
[
    {
        "$set": {
            "diff": {
                "$objectToArray": "$diff"
            }
        }
    }, {
        "$match": {
            "diff.k": {"$regex": "(^tests?)|(tests?.java$)","$options":"i"}
        }
    }, {
        "$group": {
            "_id": "$users.author.github_id", 
            "nCommitsChangingTestFile": {
                "$sum": 1
            }
        }
    }
]
')

user_changing_test_files <- user_changing_test_files %>% inner_join(users, by = "_id")
```


```{r}
p_user_changing_test_files <-  plot_users(user_changing_test_files)
save_plot(p_user_changing_test_files)
print(p_user_changing_test_files)

```





```{r}
#Get file modifications
files_modification <- commits_collection$aggregate('
[
    {
      "$match":{
          "users.author":{"$exists":1}
      }
    },
    {
        "$set": {
            "author_id": "$users.author.github_id"
        }
    }, {
        "$project": {
            "diff": 1, 
            "author_id": 1
        }
    }, {
        "$set": {
            "diff": {
                "$objectToArray": "$diff"
            }
        }
    }, {
        "$unwind": "$diff"
    }, {
        "$project": {
            "diff": 0
        }
    }
]
')

files_modification <- files_modification %>% inner_join(users, by = c('author_id' = '_id'))
```


```{r}
plot_users(files_modification)
```





```{r}
#Get test files modifications
test_files_modification <- commits_collection$aggregate('
[
    {
        "$set": {
            "author_id": "$users.author.github_id"
        }
    }, {
        "$project": {
            "users": 0, 
            "author_email": 0, 
            "author": 0, 
            "commiter": 0, 
            "commiter_email": 0, 
            "subject": 0, 
            "url": 0
        }
    }, {
        "$set": {
            "diff": {
                "$objectToArray": "$diff"
            }
        }
    }, {
        "$project": {
            "diff.v": 0
        }
    }, {
        "$match": {
            "diff.k": {
                "$regex": "(^tests?)|(tests?.java$)", 
                "$options": "i"
            }
        }
    }, {
        "$unwind": "$detection"
    }, {
        "$set": {
            "file": "$detection.TestClass"
        }
    }, {
        "$project": {
            "detection.App": 0, 
            "detection.TestClass": 0, 
            "detection.TestFilePath": 0, 
            "detection.ProductionFilePath": 0, 
            "detection.RelativeTestFilePath": 0, 
            "detection.NumberOfMethods": 0, 
            "detection.RelativeProductionFilePath": 0
        }
    }, {
        "$match": {
            "$expr": {
                "$in": [
                    "$file", "$diff.k"
                ]
            }
        }
    }, {
        "$project": {
            "diff": 0
        }
    }
]
')

test_files_modification <- test_files_modification %>% inner_join(users, by = c('author_id' = '_id'))
```


```{r}
p_test_files_modification <- plot_users(test_files_modification)
save_plot(p_test_files_modification)
p_test_files_modification
```

## RQ2.2 How does the distribution of developper introducing test smell compare to the global distribution

```{r}
#Get variation in test smells between consecutive commits
test_smells_variations <- commits_collection$aggregate('
[
    {
        "$set": {
            "author_id": "$users.author.github_id"
        }
    }, {
        "$project": {
            "users": 0, 
            "author_email": 0, 
            "author": 0, 
            "commiter": 0, 
            "commiter_email": 0, 
            "subject": 0, 
            "url": 0
        }
    }, {
        "$set": {
            "diff": {
                "$objectToArray": "$diff"
            }
        }
    }, {
        "$project": {
            "diff.v": 0
        }
    }, {
        "$match": {
            "diff.k": {
                "$regex": "(^tests?)|(tests?.java$)", 
                "$options": "i"
            }
        }
    }, {
        "$unwind": "$detection"
    }, {
        "$set": {
            "file": "$detection.TestClass"
        }
    }, {
        "$project": {
            "detection.App": 0, 
            "detection.TestClass": 0, 
            "detection.TestFilePath": 0, 
            "detection.ProductionFilePath": 0, 
            "detection.RelativeTestFilePath": 0, 
            "detection.NumberOfMethods": 0, 
            "detection.RelativeProductionFilePath": 0
        }
    }, {
        "$match": {
            "$expr": {
                "$in": [
                    "$file", "$diff.k"
                ]
            }
        }
    }, {
        "$project": {
            "diff": 0
        }
    }, {
        "$set": {
            "author_timestamp": {
                "$toLong": "$author_timestamp"
            }, 
            "commiter_timestamp": {
                "$toLong": "$commiter_timestamp"
            }
        }
    }, {
        "$setWindowFields": {
            "partitionBy": {
                "app": "$app", 
                "file": "$file"
            }, 
            "sortBy": {
                "author_timestamp": 1
            }, 
            "output": {
                "commitRank": {
                    "$rank": {}
                }
            }
        }
    }, {
        "$setWindowFields": {
            "partitionBy": {
                "app": "$app", 
                "file": "$file"
            }, 
            "sortBy": {
                "commitRank": 1
            }, 
            "output": {
                "variationAssertionRoulette": {
                    "$derivative": {
                        "input": "$detection.Assertion Roulette"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationConditionalTestLogic": {
                    "$derivative": {
                        "input": "$detection.Conditional Test Logic"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationConstructorInitialization": {
                    "$derivative": {
                        "input": "$detection.Constructor Initialization"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationDefaultTest": {
                    "$derivative": {
                        "input": "$detection.Default Test"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationEmptyTest": {
                    "$derivative": {
                        "input": "$detection.EmptyTest"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationExceptionCatchingThrowing": {
                    "$derivative": {
                        "input": "$detection.Exception Catching Throwing"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationGeneralFixture": {
                    "$derivative": {
                        "input": "$detection.General Fixture"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationMysteryGuest": {
                    "$derivative": {
                        "input": "$detection.Mystery Guest"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationPrintStatement": {
                    "$derivative": {
                        "input": "$detection.Print Statement"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationRedundantAssertion": {
                    "$derivative": {
                        "input": "$detection.Redundant Assertion"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationSensitiveEquality": {
                    "$derivative": {
                        "input": "$detection.Sensitive Equality"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationVerboseTest": {
                    "$derivative": {
                        "input": "$detection.Verbose Test"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationSleepyTest": {
                    "$derivative": {
                        "input": "$detection.Sleepy Test"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationDuplicateAssert": {
                    "$derivative": {
                        "input": "$detection.Duplicate Assert"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationUnknowTest": {
                    "$derivative": {
                        "input": "$detection.Unknown Test"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationIgnoredTest": {
                    "$derivative": {
                        "input": "$detection.IgnoredTest"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationResourceOptimism": {
                    "$derivative": {
                        "input": "$detection.Resource Optimism"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationMagicNumberTest": {
                    "$derivative": {
                        "input": "$detection.Magic Number Test"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }, 
                "variationDependentTest": {
                    "$derivative": {
                        "input": "$detection.Dependent Test"
                    }, 
                    "window": {
                        "documents": [
                            -1, 0
                        ]
                    }
                }
            }
        }
    }, {
        "$sort": {
            "app": 1, 
            "file": 1, 
            "commitRank": 1
        }
    }
]
')

#Compute global variation
test_smells_variations <- test_smells_variations %>% mutate(globalVariation = rowSums(select(., starts_with('variation'))))
test_smells_variations <- test_smells_variations %>% inner_join(users, by = c('author_id' = '_id'))
```


```{r}
#Extracts smelly commits
smelly_test_file_modification <- test_smells_variations %>% filter(globalVariation > 0)
#smelly_test_file_modification <- test_smells_variations %>% filter(if_any(starts_with('variation'), ~ . > 0))

```

```{r}
p_smelly_test_file_modification <- plot_users(smelly_test_file_modification)
save_plot(p_smelly_test_file_modification)
p_smelly_test_file_modification
```



```{r}
smelly_internal_test <- execute_test(test_files_modification$internal, smelly_test_file_modification$internal)

smelly_external_test <- execute_test(test_files_modification$external, smelly_test_file_modification$external)

print(smelly_internal_test)
print(smelly_external_test)
```
### By smells type

```{r}
smelly_modification_by_type <- test_smells_variations %>% select(starts_with('variation'), internal, external) %>% gather(key='smells', value='count' ,-internal, -external) %>% gather(key='type', value='Expertise', -smells, -count) %>% filter(count > 0) %>% group_by(type, smells, Expertise) %>% summarise(count = n()) %>% mutate(percentage = count/sum(count)) %>% mutate(smells = stringr::str_replace(smells,"^variation",""))
  
p_smelly_modification_by_type_internal <- smelly_modification_by_type %>% filter(type == 'internal') %>% plot_by_types()
  
save_plot(p_smelly_modification_by_type_internal, heigth = 17)

p_smelly_modification_by_type_external <-  smelly_modification_by_type %>% filter(type == 'external') %>% plot_by_types()


save_plot(p_smelly_modification_by_type_external, heigth = 17)

print(p_smelly_modification_by_type_internal)
print(p_smelly_modification_by_type_external)
```

## RQ2.3 How does the distribution of developper removing test smell compare to the global distribution

```{r}
improving_test_file_modification <- test_smells_variations %>% filter(globalVariation < 0)


```

```{r}
p_improving_test_file_modification <- plot_users(improving_test_file_modification)
save_plot(p_improving_test_file_modification)
print(p_improving_test_file_modification)
```

```{r}
improving_internal_test <-execute_test(test_files_modification$internal, improving_test_file_modification$internal)

improving_external_test <-execute_test(test_files_modification$external, improving_test_file_modification$external)

print(improving_external_test)
print(improving_internal_test)
```

```{r}

improving_modification_by_type <- test_smells_variations %>% select(starts_with('variation'), internal, external) %>% gather(key='smells', value='count' ,-internal, -external) %>% gather(key='type', value='Expertise', -smells, -count) %>% filter(count < 0) %>% group_by(type, smells, Expertise) %>% summarise(count = n()) %>% mutate(percentage = count/sum(count)) %>% mutate(smells = stringr::str_replace(smells,"^variation",""))
  
p_improving_modification_by_type_internal <- improving_modification_by_type %>% filter(type == 'internal') %>% plot_by_types()

save_plot(p_improving_modification_by_type_internal, heigth = 17)

p_improving_modification_by_type_external <- improving_modification_by_type %>% filter(type == 'external') %>% plot_by_types()

save_plot(p_improving_modification_by_type_external, heigth = 17)

print(p_improving_modification_by_type_internal)
print(p_improving_modification_by_type_external)
```

## Tables


```{r}
get_info <- function(data, metric, level){
  if(metric == 'internal'){
    nrow(data %>% filter(internal == level))
  } else {
    nrow(data %>% filter(external == level))
  }
}

t <- data.frame(c(get_info(test_smells_variations,'internal','Senior'), 
             get_info(smelly_test_file_modification,'internal','Senior'),
             get_info(improving_test_file_modification,'internal','Senior')),
           c(get_info(test_smells_variations,'external','Senior'),
             get_info(smelly_test_file_modification,'external','Senior'),
             get_info(improving_test_file_modification,'external','Senior')),
           c(get_info(test_smells_variations,'internal','Junior'),
             get_info(smelly_test_file_modification,'internal','Junior'),
             get_info(improving_test_file_modification,'internal','Junior')),
           c(get_info(test_smells_variations,'external','Junior'),
             get_info(smelly_test_file_modification,'external','Junior'),
             get_info(improving_test_file_modification,'external','Junior')))
t$name <- c('All', 'Smelly', 'Refactoring')
colnames(t) <- c('Internal|Senior', 'External|Senior','Internal|Junior' ,'External|Junior', 'Test file changes')

table <- t %>% flextable(col_keys = c('Test file changes', 'Internal|Senior', 'Internal|Junior', 'External|Senior', 'External|Junior')) %>% colformat_int(big.mark = ' ')
table
table %>% save_as_image('images/summary.pdf', webshot = "webshot")

```